### 2017-11-14 paper/sections/1_introduction.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

In this paper, we design cache management policies for ParSplice, driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions.

The memory usage for a ParSplice cache that stores molecule coordinates is
shown in Figure~\ref{fig:cache-management}. The solid lines are the existing
cache management policies, exposed as configurations in ParSplice.  The default
ParSplice configuration uses an unlimited sized cache, shown by the ``No Cache
Management" line, but using this much memory for one cache is unacceptable for
HPC environments, where memory is precious and a common goal is to keep memory
for such data structures below 3\%\footnote{Empirically, we find this threshold
works well for most applications}.  While users can configure ParSplice to
resize the cache when it reaches a certain threshold, this solution requires
tuning and parameter sweeps to find the optimal cache size that sacrifices
negligible performance.  Furthermore, this threshold changes with different
initial configurations and clusters, as shown by our keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}.  The ``Cache (too small)"
curve in Figure~\ref{fig:cache-management} shows how a poorly configured cache
can save memory but at the expense of performance.  Our dynamically sized
cache, shown by the dashed line in Figure~\ref{fig:cache-management} detects
keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without sacrificing performance.  More importantly, it works for a variety
initial conditions without changing the policy itself.

% What is Mantle
To design more flexible cache management policies, like our dynamically sized
cache solution in Figure~\ref{fig:cache-management}, we use the data management
language and policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.
This framework allows us to dynamically explore the effects of different
software-defined cache management strategies for the changing key-value
workloads generated by ParSplice.  Mantle was originally touted as a
programmable file system metadata load balancer, but we realize now that the
collection of abstractions designed for file systems was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service and show how this approach is useful for reasoning about
and designing different cache management strategies in ParSplice.  By service,
we mean a system that manages data and responds to requests, such as a file
system or key-value store.  Developers write policies for ``when" they want
data moved and ``how much" of the data to move, then the framework executes
these policies whenever a decision needs to be made.  These abstractions help
developers unfamiliar with the domain quickly reason about, develop, and deploy
new policies that control temporal and spatial locality. We show that Mantle:

>>>>

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
the existing cache management policies, exposed as configurations in ParSplice.
The default ParSplice configuration uses an unlimited sized cache, shown by the
``No Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where memory is precious and a common goal
is to keep memory for such data structures below 3\%\footnote{Empirically, we
find this threshold works well for most applications}.  While users can
configure ParSplice to resize the cache when it reaches a certain threshold,
this solution requires tuning and parameter sweeps to find the optimal cache
size that sacrifices negligible performance.  Furthermore, this threshold
changes with different initial configurations and clusters, as shown by our
keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the expense of performance.  Our
dynamically sized cache, shown by the dashed line in
Figure~\ref{fig:cache-management}, detects keyspace access patterns and
re-sizes the cache accordingly. The design of this policy is driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions.  Without tuning or parameter
sweeps, our solution saves more memory than a hand-tuned cache without
sacrificing performance.  More importantly, it works for a variety initial
conditions without changing the policy itself.

% What is Mantle
To design more flexible cache management policies, like our dynamically sized
cache solution in Figure~\ref{fig:cache-management}, we use the data management
language and policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.
This framework allows us to dynamically explore the effects of different
software-defined cache management strategies for the changing key-value
workloads generated by ParSplice.  Mantle was originally touted as a
programmable file system metadata load balancer, but we realize now that the
collection of abstractions designed for file systems was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service, such as a file system or key-value store, and show how
this approach is useful for reasoning about and designing different cache
management strategies in ParSplice.  Developers write policies for ``when" they
want data moved and ``how much" of the data to move, then the framework
executes these policies whenever a decision needs to be made.  These
abstractions help developers unfamiliar with the domain quickly deploy new
policies {\it or} well-understood techniques from other domains.  We show that
Mantle:

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The actual policy is shown in Figure~\ref{src:lru-dyn} where the only
difference between the code snippet in Figure~\ref{src:lru} is the addition of
lines 2-7. Note that policy also uses the \texttt{RDState} and \texttt{WRState}
functions described in Section~\S\ref{sec:extracting}.

\begin{figure}[t]
\footnotesize
\centering
\begin{minted}[xleftmargin=3em,linenos]{lua}
function when()
  if server[whoami]['cachesize'] > n then
    if server[whoami]['cachesize'] > 100K then
      WRstate(1)
    end
    if RDstate() == 1 then
      return true
    end
  end
  return false
end
\end{minted}
\caption{A more complicated LRU-based cache policy that has two parts. After
absorbing the initial burstiness of the workload, the policy switches to a more
constrained cache to limit memory usage. The performance/utilization for
different values of \texttt{n} is shown in
Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
\end{figure}

We trigger the policy switch at 100K keys to absorb the flash crowd at the
beginning of the run. Once triggered, keys are evicted to bring the size of the
cache down to the threshold.  In the bar chart, the cache sizes for the \(n\)
key limit policy are along the \(x\) axis.

>>>>

The actual policy is shown and described in more detail in the next section in
Figure~\ref{src:lru-dyn}, but the only difference between the code snippet in
Figure~\ref{src:lru} is the addition of lines 2-7.  We trigger the policy
switch at 100K keys to absorb the flash crowd at the beginning of the run. Once
triggered, keys are evicted to bring the size of the cache down to the
threshold.  In the bar chart, the cache sizes for the \(n\) key limit policy
are along the \(x\) axis.

### 2017-11-14 paper/sections/5_file-systems.tex
paper: Reza comments - intro: address cache management as system challenge in ParSplice - intro: this lets us deploy existing techniques from other domains - intro: clarify that Fig1 shows general problem (with preview of solution) - scope: pseudo code side by side
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@

\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
Fig~\ref{src:lua-cephfs}: Lines 3,4,10) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded.

@@

\begin{figure}[t]
\footnotesize
\begin{minted}[xleftmargin=3em,linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    overloaded = RDstate() + 1
    WRstate(overloaded)
    if overloaded > 2 then
      return true
    end
  end
  else then
    WRstate(0)
  end
  return false
end
\end{minted}
\caption{CephFS file system metadata load balancer.\label{src:lua-cephfs}}
\end{figure}

>>>>

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \footnotesize
        \centering
        \begin{minted}[xleftmargin=3em,linenos]{lua}
function when()
  if server[whoami]['cachesize'] > n then
    if server[whoami]['cachesize'] > 100K then
      WRstate(1)
    end
    if RDstate() == 1 then
      return true
    end
  end
  return false
end
        \end{minted}
        \caption{ParSplice cache management policy that absorbs the burstiness of the
        workload before switching to a constrained cache.  The
        performance/utilization for different  \texttt{n} is in
        Figure~\ref{fig:methodology-tradeoff-dynamic}. \label{src:lru-dyn}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \footnotesize
        \begin{minted}[xleftmargin=3em,linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    overloaded = RDstate() + 1
    WRstate(overloaded)
    if overloaded > 2 then
      return true
    end
  end
  else then WRstate(0) end
  return false
end
        \end{minted}
	\caption{CephFS file system metadata load balancer, designed in 2004
        in~\cite{weil:sc2004-dyn-metadata}, reimplemented in Lua
        in~\cite{sevilla:sc15-mantle}. This policy has many similarities to the
        ParSplice cache management policy.\label{src:lua-cephfs}}
    \end{subfigure}
    \caption{ParSplice's cache management policy has the same components as
    CephFS's load balancing policy.}
\end{figure*}

@@

\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
Fig~\ref{src:lua-cephfs}: Lines 3,4,9) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded.

@@
